### 2014-06-18 ApTranscriptome_TR.Rmd
Updates to GSEA clustering and mosaic plot

# Reorganize for plotting to show overlap among categories between colonies
type.table <- table(Ap.response.type[ , 'Ar.type'], Ap.response.type[ , 'A22.type'])
# reorder 
tt2 <- type.table[c(2,4,1,3,5), c(2,4,1,3,5)]

# plot
mosaic.colors <- brewer.pal(5, "Blues")
par(mar = c(5.1, 3.1, 2.1, 7.5))
mosaicplot(tt2, ylab = expression(italic("A. picea")), xlab = "", main = "", color = mosaic.colors, cex.axis = 0.01)
labels <- c("High", "Low", "Bimodal", "Intermediate", "NotResp")
text(c(.1,.3,.55,.75,.95), par("usr")[3] - 0.05, srt = 45, adj = 1, labels = labels, xpd = TRUE, font = 1)
text(.5, par("usr")[4], "A. carolinensis", xpd = TRUE, font = 3)
legend(1.05, .75, fill = mosaic.colors, c("High", "Low", "Bimodal", "Intermediate", "NotResp"), xpd=TRUE)
```

@@

maxexplot <- ggplot(Ap.df, aes(x=max.val, fill=colony)) + 
  geom_density(alpha=0.2, position="identity") + 
  #scale_fill_manual(name = "Colony", values = c("white", "black")) + 
  scale_y_continuous(name="Density") +
  scale_x_continuous(name="Temperature of maximum expression")
print(maxexplot)
```

@@

# plot
T_on_plot <- ggplot(Ap.T_on, aes(x = T_on, y = ..density.., colour = colony, group = colony, fill = colony)) +
    geom_histogram(position = "dodge", binwidth = 3) +
    geom_density(adjust = 3, fill = NA, size = 2) + 
    scale_y_continuous(name="Density") +
    scale_x_continuous(name=expression("Temperature"))
print(T_on_plot)
```

@@

# report items in each cluster at height 0.6 
GSEAReportClusters(Ap.high.hclust, h = 0.6)
```

@@

# report items in each cluster at height 2
GSEAReportClusters(Ap.low.hclust, h = 2)
```

@@

# report items in each cluster at height 1.5
GSEAReportClusters(Ap.bim.hclust, h = 1.5)
```

```{r gsea_Ap_int, results = 'hide'}
## Genes with *Intermediate* expression in both colonies
Ap.int <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate" & Ap.response.type$Ar.type == "Intermediate"), "Transcript"]
# create gene list, setting value to 1 for "int" transcripts
Ap.geneList.int <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.int) <- names(Ap.geneList)
Ap.geneList.int[(which(names(Ap.geneList.int) %in% Ap.int))] <- 1
# check correct number of values set to 1
table(Ap.geneList.int)
# Run GSEA
Ap.int.gsea <- gsea(genelist = Ap.geneList.int, geneID2GO = geneID2GO)
```

@@

# report items in each cluster at height 1.5
GSEAReportClusters(Ap.int.hclust, h = 1.5)
```

## merge results into single table
Ap.high.gsea$Type <- "High"
Ap.low.gsea$Type <- "Low"
Ap.bim.gsea$Type <- "Bimodal"
Ap.int.gsea$Type <- "Intermediate"
# combine
Ap.gsea <- rbind(Ap.high.gsea, Ap.low.gsea, Ap.bim.gsea, Ap.int.gsea)
# reorder
Ap.gsea1 <- Ap.gsea[,c("Type", "GO.ID", "Term", "Annotated", "Significant", "Expected", "parentchild")]
colnames(Ap.gsea1)[7] <- "P"
# write to file
write.csv(Ap.setdiff.gsea, file = paste(resultsdir, "Ap_setdiff_GSEA_", Sys.Date(), ".csv", sep = ""), row.names = FALSE)
```

```{r gsea_by_type_table, echo = FALSE, results = 'asis'}
#pandoc.table(Ap.gsea1[, 1:3], caption = "Enriched GO terms for each thermal-response category")
print(xtable(Ap.gsea1[ , 1:3], comment = getOption("xtable.comment", FALSE)))
```

@@

# report items in each cluster at height 1.5
GSEAReportClusters(Ar.int.A22.bim.gsea.term.sim.hclust, h = 1)
```

@@

```{r gsea_Ar_int_A22_low_cluster}
# similarity among enriched GO terms
Ar.int.A22.low.gsea.term.sim <- termSim(Ar.int.A22.low.gsea$GO.ID, Ar.int.A22.low.gsea$GO.ID, ont = "BP", organism = 'fly')
# replace GO IDs with terms
rownames(Ar.int.A22.low.gsea.term.sim) <- Ar.int.A22.low.gsea$Term
colnames(Ar.int.A22.low.gsea.term.sim) <- Ar.int.A22.low.gsea$Term
# distance matrix
Ar.int.A22.low.gsea.term.sim.hclust <- hclust(dist(Ar.int.A22.low.gsea.term.sim))
plot(Ar.int.A22.low.gsea.term.sim.hclust)
# report items in each cluster at height 1.8
GSEAReportClusters(Ar.int.A22.low.gsea.term.sim.hclust, h = 1.6)
```

@@

```{r gsea_A22_high_cluster}
# similarity among enriched GO terms
A22.high.gsea.term.sim <- termSim(A22.high.gsea$GO.ID, A22.high.gsea$GO.ID, ont = "BP", organism = 'fly')
# replace GO IDs with terms
rownames(A22.high.gsea.term.sim) <- A22.high.gsea$Term
colnames(A22.high.gsea.term.sim) <- A22.high.gsea$Term
# distance matrix
A22.high.gsea.term.sim.hclust <- hclust(dist((A22.high.gsea.term.sim)))
plot((A22.high.gsea.term.sim.hclust))
# report items in each cluster at height 1.8
GSEAReportClusters(A22.high.gsea.term.sim.hclust, h = 1.2)
```
GO enrichment for *A. picea* **Low** genes.

@@

@@

@@

ggplot(resp.TPM.dt.sub.pred["100348|*|comp3450455_c0_seq1"], aes(x=val, y=pTPM.scaled, group=Transcript)) +
  geom_line() + 
  facet_grid(. ~ colony) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))

>>>>

# Reorganize for plotting to show overlap among categories between colonies
type.table <- table(Ap.response.type[ , 'Ar.type'], Ap.response.type[ , 'A22.type'])
# reorder
tt2 <- type.table[c("Low", "Intermediate", "High", "Bimodal", "NotResp"), c("Low", "Intermediate", "High", "Bimodal", "NotResp")]

# plot
png("mosaic_plot.png")
mosaic.colors <- brewer.pal(5, "Blues")
par(mar = c(5.1, 3.1, 2.1, 7.5))
mosaicplot(tt2, ylab = expression(italic("A. picea")), xlab = "", main = "", color = mosaic.colors, cex.axis = 0.01)
labels <- c("Low", "Intermediate", "High", "Bimodal", "NotResp")
text(c(.1,.5,.75,.85,.99), par("usr")[3] - 0.05, srt = 45, adj = 1, labels = labels, xpd = TRUE, font = 1)
text(.5, par("usr")[4], "A. carolinensis", xpd = TRUE, font = 3)
legend(1.05, .75, fill = mosaic.colors, c("Low", "Intermediate", "High", "Bimodal", "NotResp"), xpd=TRUE)
dev.off()
```

@@

maxexplot <- ggplot(Ap.df, aes(x=max.val, fill=colony)) + 
  geom_density(alpha=0.2, position="identity") + 
  labs(x = "Temperature of maximum expression", y = "Density") +
  theme(axis.title = element_text(size = rel(2))) +
  theme(axis.text = element_text(size = rel(1.2)))
print(maxexplot)

# Figure for presentation
png("results/temp_max_expression.png")
maxexplot + theme(legend.position = "none")
dev.off()
```

@@

# low plot
png("T_on_low.png")
T_on_low_plot <- ggplot(rbind(A22.low.T_on, Ar.low.T_on), aes(x = T_on, y = ..density.., colour = colony, group = colony, fill = colony)) +
    geom_histogram(position = "dodge", binwidth = 3) +
    geom_density(adjust = 3, fill = NA, size = 2) +
    labs(x = "Temperature of gene activation", y = "Density") +
    theme(axis.title = element_text(size = rel(2))) +
    theme(axis.text = element_text(size = rel(1.2))) +
    theme(legend.position = "none")
print(T_on_low_plot)
dev.off()

# high plot
png("T_on_high.png")
T_on_high_plot <- ggplot(rbind(A22.high.T_on, Ar.high.T_on), aes(x = T_on, y = ..density.., colour = colony, group = colony, fill = colony)) +
    geom_histogram(position = "dodge", binwidth = 3) +
    geom_density(adjust = 3, fill = NA, size = 2) +
    labs(x = "Temperature of gene activation", y = "Density") +
    theme(axis.title = element_text(size = rel(2))) +
    theme(axis.text = element_text(size = rel(1.2))) +
    theme(legend.position = "none")
print(T_on_high_plot)
dev.off()

# combined plot
T_on_plot <- ggplot(Ap.T_on, aes(x = T_on, y = ..density.., colour = colony, group = colony, fill = colony)) +
    geom_histogram(position = "dodge", binwidth = 3) +
    geom_density(adjust = 3, fill = NA, size = 2) + 
    labs(x = "Temperature of gene activation", y = "Density") +
    theme(axis.title = element_text(size = rel(2))) +
    theme(axis.text = element_text(size = rel(1.2))) +
    theme(legend.position = "none")
print(T_on_plot)
```

@@

# report items in each cluster
GSEAReportClusters(Ap.high.hclust, h = 1)
write.table(unlist(GSEAReportClusters(Ap.high.hclust, h = 0.6)), file = "Ap_high_hclust.txt")
```

@@

# report items in each cluster at height 2
(Ap.low.hclust.report <- GSEAReportClusters(Ap.low.hclust, h = 2))
```

@@

# report items in each cluster at height 1.5
(Ap.bim.hclust.report <- GSEAReportClusters(Ap.bim.hclust, h = 1.5))
```

```{r gsea_Ap_int, results = 'hide'}
# genes with *Intermediate* expression in both colonies
Ap.int <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate" & Ap.response.type$Ar.type == "Intermediate"), "Transcript"]
# create gene list, setting value to 1 for "int" transcripts
Ap.geneList.int <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.int) <- names(Ap.geneList)
Ap.geneList.int[(which(names(Ap.geneList.int) %in% Ap.int))] <- 1
# check correct number of values set to 1
table(Ap.geneList.int)
# run GSEA
Ap.int.gsea <- gsea(genelist = Ap.geneList.int, geneID2GO = geneID2GO)
```

@@

# report items in each cluster at height 1.5
(Ap.int.hclust.report <- GSEAReportClusters(Ap.int.hclust, h = 1.5))
```

@@

# report items in each cluster at height 1.5
(Ar.int.A22.bim.hclust.report <- GSEAReportClusters(Ar.int.A22.bim.gsea.term.sim.hclust, h = 1))
```

@@

```{r gsea_Ar_int_A22_low_cluster}
# similarity among enriched GO terms
Ar.int.A22.low.gsea.term.sim <- termSim(Ar.int.A22.low.gsea$GO.ID, Ar.int.A22.low.gsea$GO.ID, ont = "BP", organism = 'fly')
# replace GO IDs with terms
rownames(Ar.int.A22.low.gsea.term.sim) <- Ar.int.A22.low.gsea$Term
colnames(Ar.int.A22.low.gsea.term.sim) <- Ar.int.A22.low.gsea$Term
# distance matrix
Ar.int.A22.low.gsea.term.sim.hclust <- hclust(dist(Ar.int.A22.low.gsea.term.sim))
plot(Ar.int.A22.low.gsea.term.sim.hclust)
# report items in each cluster at height 1.8
(Ar.int.A22.low.hclust.report <- GSEAReportClusters(Ar.int.A22.low.gsea.term.sim.hclust, h = 1.6))
```

@@

```{r gsea_A22_high_cluster}
# similarity among enriched GO terms
A22.high.gsea.term.sim <- termSim(A22.high.gsea$GO.ID, A22.high.gsea$GO.ID, ont = "BP", organism = 'fly')
# replace GO IDs with terms
rownames(A22.high.gsea.term.sim) <- A22.high.gsea$Term
colnames(A22.high.gsea.term.sim) <- A22.high.gsea$Term
# distance matrix
A22.high.gsea.term.sim.hclust <- hclust(dist((A22.high.gsea.term.sim)))
plot((A22.high.gsea.term.sim.hclust))
# report items in each cluster at height 1.8
GSEAReportClusters(A22.high.gsea.term.sim.hclust, h = 1.2)
```

GO enrichment for *A. picea* **Low** genes.

@@

Clustering of enriched GO terms for *A. picea* **Bimodal** genes.

@@

GO enrichment for *A. carolinensis* **High** genes.

```{r gsea_Ar_high, results = 'hide'}
# Ar 'High' genes
Ar.geneList.high <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.high) <- names(Ap.geneList)
Ar.geneList.high[(which(names(Ar.geneList.high) %in% Ar.high.transcripts))] <- 1
Ar.high.gsea <- gsea(genelist = Ar.geneList.high, geneID2GO = geneID2GO)
```

Clustering of enriched GO terms for *A. carolinensis* **High** genes.

```{r gsea_Ar_high_cluster}
# similarity among enriched GO terms
Ar.high.gsea.term.sim <- termSim(Ar.high.gsea$GO.ID, Ar.high.gsea$GO.ID, ont = "BP", organism = 'fly')
# replace GO IDs with terms
rownames(Ar.high.gsea.term.sim) <- Ar.high.gsea$Term
colnames(Ar.high.gsea.term.sim) <- Ar.high.gsea$Term
# distance matrix
Ar.high.gsea.term.sim.hclust <- hclust(dist((Ar.high.gsea.term.sim)))
plot((Ar.high.gsea.term.sim.hclust))
# report items in each cluster
GSEAReportClusters(Ar.high.gsea.term.sim.hclust, h = 1.2)
```
GO enrichment for *A. carolinensis* **Low** genes.

```{r gsea_Ar_low, results = 'hide'}
# Ar 'Low' genes
Ar.geneList.low <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.low) <- names(Ap.geneList)
Ar.geneList.low[(which(names(Ar.geneList.low) %in% Ar.low.transcripts))] <- 1
Ar.low.gsea <- gsea(genelist = Ar.geneList.low, geneID2GO = geneID2GO)
```

Clustering of enriched GO terms for *A. carolinensis* **Low** genes.

```{r gsea_Ar_low_cluster}
# similarity among enriched GO terms
Ar.low.gsea.term.sim <- termSim(Ar.low.gsea$GO.ID, Ar.low.gsea$GO.ID, ont = "BP", organism = 'fly')
# replace GO IDs with terms
rownames(Ar.low.gsea.term.sim) <- Ar.low.gsea$Term
colnames(Ar.low.gsea.term.sim) <- Ar.low.gsea$Term
# distance matrix
Ar.low.gsea.term.sim.hclust <- hclust(dist((Ar.low.gsea.term.sim)))
plot((Ar.low.gsea.term.sim.hclust))
# report items in each cluster at height 1.8
GSEAReportClusters(Ar.low.gsea.term.sim.hclust, h = 2)
```

GO enrichment for *A. carolinensis* **Bimodal** genes.

```{r gsea_Ar_bim, results = 'hide'}
# Ar 'Bim' genes
Ar.geneList.bim <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.bim) <- names(Ap.geneList)
Ar.geneList.bim[(which(names(Ar.geneList.bim) %in% Ar.bim.transcripts))] <- 1
Ar.bim.gsea <- gsea(genelist = Ar.geneList.bim, geneID2GO = geneID2GO)
```

Clustering of enriched GO terms for *A. carolinensis* **Bimodal** genes.

```{r gsea_Ar_bim_cluster}
# similarity among enriched GO terms
Ar.bim.gsea.term.sim <- termSim(Ar.bim.gsea$GO.ID, Ar.bim.gsea$GO.ID, ont = "BP", organism = 'fly')
# replace GO IDs with terms
rownames(Ar.bim.gsea.term.sim) <- Ar.bim.gsea$Term
colnames(Ar.bim.gsea.term.sim) <- Ar.bim.gsea$Term
# distance matrix
Ar.bim.gsea.term.sim.hclust <- hclust(dist((Ar.bim.gsea.term.sim)))
plot((Ar.bim.gsea.term.sim.hclust))
# report items in each cluster at height 1.8
GSEAReportClusters(Ar.bim.gsea.term.sim.hclust, h = 1)
```

GO enrichment for *A. carolinensis* **Intermediate** genes.

```{r gsea_Ar_int, results = 'hide'}
# Ar 'Int' genes
Ar.geneList.int <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.int) <- names(Ap.geneList)
Ar.geneList.int[(which(names(Ar.geneList.int) %in% Ar.int.transcripts))] <- 1
Ar.int.gsea <- gsea(genelist = Ar.geneList.int, geneID2GO = geneID2GO)
```

Clustering of enriched GO terms for *A. carolinensis* **Intermediate** genes.

```{r gsea_Ar_int_cluster}
# similarity among enriched GO terms
Ar.int.gsea.term.sim <- termSim(Ar.int.gsea$GO.ID, Ar.int.gsea$GO.ID, ont = "BP", organism = 'fly')
# replace GO IDs with terms
rownames(Ar.int.gsea.term.sim) <- Ar.int.gsea$Term
colnames(Ar.int.gsea.term.sim) <- Ar.int.gsea$Term
# distance matrix
Ar.int.gsea.term.sim.hclust <- hclust(dist((Ar.int.gsea.term.sim)))
plot((Ar.int.gsea.term.sim.hclust))
# report items in each cluster
GSEAReportClusters(Ar.int.gsea.term.sim.hclust, h = 1.8)
```

@@

png("RxN_example.png")
ggplot(resp.TPM.dt.sub.pred["100348|*|comp3450455_c0_seq1"], aes(x=val, y=pTPM.scaled, group=Transcript, color = Transcript)) +
  geom_line() + 
  facet_grid(. ~ colony) +
  labs(y = "Expression (scaled)", x = expression("Temperature ", degree, "C")) +
  theme(legend.position = "none")
dev.off()
